The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-08-24 manuscript/rebuttal.Rmd

## Reviewer 3

### R3 Point 1

> The paper is mostly well written and addresses an important topic: computational reproducibility and
> tools to achieve it. The authors describe five "pillars" under which various tools and techniques can
> be categorized. The paper is pretty comprehensive in what it covers and has useful supplementary
> material. My main issue with the paper is that it reiterates many things that have been covered in
> other papers. The authors cite 11 such papers. Many of the topics covered in submitted paper have
> already been covered well in other papers. I am most familiar with reference 13, which covers many of
> the same topics, although this paper goes into more detail on many issues. The authors could do more
> to differentiate their paper from previous ones and perhaps remove some topics that are already
> covered well elsewhere.

RESPONSE: We thank reviewer 3 for their careful assessment of our article and appreciate the suggestions
which have improved the manuscript.
Indeed, this article does reiterate a great deal of what has already been said, it is a review after
all.
We note that each of those recommendation articles mentioned have different emphases.
For example, some of them don't speak to the benefits of literate programming,
while others do not mention the importance of documentation.
Some are focused on specific concepts like infrastructure, containerisation or
continuous analysis.
Some are proof of principle workflows, which present a rigid combination of tools.
Here, we describe the ways computational researchers can make their own work more reproducible.
In addition to the Five Pillars infographic, the specific combination of recommendations
is unique.
And ours is the only such article that provides a handy list of recommended tutorials as a supplementary
resource.

### 2023-08-24 manuscript/5pillars.Rmd
rev3 again

A number of guides recommending enhanced computational reproducibility practices have been developed
[ @Tan2010-is; @Peng2011-ia; @Sandve2013-lk; @Piccolo2016-ji; @Lewis2016-qh;
@Gruning2018-sg; @Akalin2018-eg; @Brito2020-ld; @Peikert2021-eq;
@Rodrigues2023-fw; @DuPre2022-bk; @Krafczyk2021-xv; @Stodden2016-aj].
Inspired by these principles, this article presents the
*five pillars of reproducible computational research* (**Figure 1**).
There are clear parallels with the established three pillars of the open science framework,
(open data, code and papers) [@Gorgolewski2016-hv].
Here, the emphasis is on practical reproducibility with an increased focus on programming practices,
transparent reporting and the role of computational environments.
The intended audience is anyone who analyzes biological data (not tool developers), however the
principles described here could equally apply to other domains of study.

```{r,5pillars,fig.cap="Figure 1. The five pillars of computational reproducibility.",echo=FALSE,fig.height=5}

@@

text( c(0.9,2.9,4.9,6.9,8.9) , c(4.5,2.5,4.5,2.5,4.5) , lab=labs)
text( 5 , 0.2 , lab="End-to-end automated process",cex=1.2)
text(5,7,lab="Code testing",cex=1.2)

@@

```{r,dvcs,fig.cap="Figure 3. Distributed version control. Adapted from @Chacon2014-ys.",echo=FALSE,fig.height=8}

@@

plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(0, 10), ylim=c(-5, 8.5))
polygon(c(-0.6,3,3,-0.6),c(0.8,0.8,5.2,5.2), col=cols2[1], border = NA)
polygon(c(-0.2,2.8,2.8,-0.2),c(1,1,2,2), col=cols[1], border = NA)
polygon(c(-0.2,2.8,2.8,-0.2),c(4,4,5,5), col=cols[1], border = NA)

@@

polygon(c(2.5,7.3,7.3,2.5),c(7,7,8.5,8.5), col=cols[4], border = NA)

labs <- c("Developer 1",
  "Working Files",
  "Local Repository",
  "Developer 2",
  "Working Files",
  "Local Repository",
  "Developer 3",
  "Working Files",
  "Local Repository")

text( c(1.2,1.2,1.2,4.9,4.9,4.9,8.5,8.5,8.5), c(0,1.5,4.5,0,1.5,4.5,0,1.5,4.5), labels=labs )
text(4.9,7.7,labels="Main Repository",cex=1.5)

arrows( 1, 2, 1, 4 ,lwd=3)
arrows( 1.4, 4, 1.4, 2 ,lwd=3)
arrows( 4.7, 2, 4.7, 4 ,lwd=3)
arrows( 5.1, 4, 5.1, 2 ,lwd=3)
arrows( 8.3, 2, 8.3, 4 ,lwd=3)
arrows( 8.7, 4, 8.7, 2 ,lwd=3)
arrows( 1, 5, 2.5, 7 ,lwd=3)
arrows( 2.9, 7 , 1.4, 5 ,lwd=3)
arrows( 4.7, 5, 4.7, 7 ,lwd=3)
arrows( 5.1, 7 , 5.1, 5 ,lwd=3)
arrows( 7.3, 7, 8.7, 5 ,lwd=3)
arrows( 8.3, 5, 6.9, 7 ,lwd=3)

text(c(0.7,4.4,8) ,c(3,3,3), labels="commit",srt=90)
text(c(1.7,5.4,9) ,c(3,3,3), labels="update",srt=-90)
text(c(1.5) ,c(6.2), labels="push",srt=50)
text(c(2.4) ,c(5.9), labels="pull",srt=50)
text(c(4.4) ,c(6), labels="push",srt=90)
text(c(5.4) ,c(6), labels="pull",srt=-90)
text(c(7.3) ,c(5.9), labels="push",srt=-60)
text(c(8.2) ,c(6.2), labels="pull",srt=-60)

mypng = readPNG('../img/version_control.png')
grid.raster(mypng, x=.55, y=.2, width=0.85)

@@

```{r,containers,fig.cap="Figure 4. Architecture of containers (left) compared to VMs (right). Adapted from @noauthor_2021-lt.",echo=FALSE,fig.height=6}

library("RColorBrewer")
par(mar = c(0.1,0.1,0.1,0.1))
cols <- brewer.pal(5, "Set2")
cols2 <- adjustcolor( cols, alpha.f = 0.7)
plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(0, 12), ylim=c(0, 8))

@@

lines(c(6,6,11.5,11.5),c(6.2,6.5,6.5,6.2)  )
text(8.5,6.7,labels="Virtualized applications")

>>>>

A number of guides recommending enhanced computational reproducibility practices have been developed
[ @Tan2010-is; @Peng2011-ia; @Sandve2013-lk; @Piccolo2016-ji; @Lewis2016-qh;
@Stodden2016-aj; @Beaulieu-Jones2017-vj; @Gruning2018-sg; @Akalin2018-eg;
@Brito2020-ld; @Peikert2021-eq; @Rodrigues2023-fw; @DuPre2022-bk; @Krafczyk2021-xv].
Inspired by these principles, we presents the *five pillars of reproducible computational research*
(**Figure 1**).
Here, the emphasis is on practical reproducibility with an increased focus on programming practices,
transparent reporting and the role of computational environments.
The intended audience is anyone who analyzes biological data (not tool developers), however the
principles described here could equally apply to other domains of study.
There are clear parallels with the established three pillars of the open science framework,
(open data, code and papers) [@Gorgolewski2016-hv].

```{r,5pillars,fig.cap="Figure 1. The five pillars of computational reproducibility.",echo=FALSE,fig.height=4.5,fig.width=6}

@@

text( c(0.9,2.9,4.9,6.9,8.9) , c(4.5,2.5,4.5,2.5,4.5) , lab=labs)
text( 5 , 0.2 , lab="End-to-end automated process",cex=1.2)
text(5,7,lab="Process validation & code testing",cex=1.2)

@@

```{r,dvcs,fig.cap="Figure 3. Distributed version control. (A) Each author has a copy of the repository where their contributions are committed before being pushed to the main central repository. Adapted from @Chacon2014-ys. (B) An example hypothetical git workflow history for a research project involving a team of three authors. Circles represent code commits. Path divergences create separate branches for independent development. Horizontal paths indicate code changes for a particular branch. Path convergences indicate where branch specific differences are incorporated into the main branch. From @ram2013.",echo=FALSE,fig.height=5.2}

suppressMessages({

@@

plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(-3.5, 13.5), ylim=c(-3.5, 7.5))
polygon(c(-0.6,3,3,-0.6),c(0.8,0.8,5.2,5.2), col=cols2[1], border = NA)
polygon(c(-0.2,2.8,2.8,-0.2),c(1,1,2,2), col=cols[1], border = NA)
polygon(c(-0.2,2.8,2.8,-0.2),c(4,4,5,5), col=cols[1], border = NA)

@@

polygon(c(2.5,7.3,7.3,2.5),c(7,7,7.7,7.7), col=cols[4], border = NA)

labs <- c("Author 1",
  "Working Files",
  "Local Repository",
  "Author 2",
  "Working Files",
  "Local Repository",
  "Author 3",
  "Working Files",
  "Local Repository")

text( c(1.3,1.3,1.3,4.9,4.9,4.9,8.5,8.5,8.5), c(0.5,1.5,4.5,0.5,1.5,4.5,0.5,1.5,4.5), labels=labs )
text(4.9,7.3,labels="Main Repository",cex=1.3)

library("shape")
Arrows( 1, 2, 1, 4 ,lwd=3, arr.type="simple")
Arrows( 1.4, 4, 1.4, 2 ,lwd=3,arr.type="simple")
Arrows( 4.7, 2, 4.7, 4 ,lwd=3, arr.type="simple")
Arrows( 5.1, 4, 5.1, 2 ,lwd=3, arr.type="simple")
Arrows( 8.3, 2, 8.3, 4 ,lwd=3, arr.type="simple")
Arrows( 8.7, 4, 8.7, 2 ,lwd=3, arr.type="simple")
Arrows( 1, 5, 2.5, 7 ,lwd=3, arr.type="simple")
Arrows( 2.9, 7 , 1.4, 5 ,lwd=3, arr.type="simple")
Arrows( 4.7, 5, 4.7, 7 ,lwd=3, arr.type="simple")
Arrows( 5.1, 7 , 5.1, 5 ,lwd=3, arr.type="simple")
Arrows( 7.3, 7, 8.7, 5 ,lwd=3, arr.type="simple")
Arrows( 8.3, 5, 6.9, 7 ,lwd=3, arr.type="simple")

text(c(0.7,4.4,8) ,c(3,3,3), labels="commit",srt=90)
text(c(1.7,5.4,9) ,c(3,3,3), labels="update",srt=-90)
text(c(1.5) ,c(6.2), labels="push",srt=45)
text(c(2.4) ,c(5.9), labels="pull",srt=45)
text(c(4.4) ,c(6), labels="push",srt=90)
text(c(5.4) ,c(6), labels="pull",srt=-90)
text(c(7.3) ,c(5.9), labels="push",srt=-55)
text(c(8.2) ,c(6.2), labels="pull",srt=-55)

text(c(1) ,c(7.5), labels="A",cex=1.7)
text(c(-3.5) ,c(0.5), labels="B",cex=1.7)

mypng = readPNG('../img/version_control.png')
grid.raster(mypng, x=.5, y=.16, width=1)

})

@@

```{r,containers,fig.cap="Figure 4. Architecture of containers (left) compared to VMs (right). Adapted from @noauthor_2021-lt.",echo=FALSE,fig.height=5}

library("RColorBrewer")
par(mar = c(0.1,0.1,0.1,0.1))
cols <- brewer.pal(5, "Set2")
cols2 <- adjustcolor( cols, alpha.f = 0.7)
plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(0, 11.5), ylim=c(0, 7))

@@

lines(c(6,6,11.5,11.5),c(6.2,6.5,6.5,6.2)  )
text(9,6.7,labels="Virtualized applications")
